---
title: "Task 1 - Palmetto BLR"
author: "Deanna Elliott"
date: "2/1/2022"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
```

### Overview

WORDS

### Data Visualization

Read In Data

```{r}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  select(species, height:green_lvs) %>% 
  mutate(species = case_when(
    species == "1" ~ "Serenoa repens",
    species == "2" ~ "Sabal etonia")) %>% 
  mutate(species = as.factor(species)) %>% 
  drop_na()
```

```{r}
# pairs plot with GGally to see overall trends

pairplot <- palmetto %>% 
  ggpairs(aes(color = species)) # canopy length and green leaves seem most                                       differentiated by species
```

#### Green Leaves vs Canopy Length by Species

```{r}

ggplot(data = palmetto,
       aes(x = length, y = green_lvs)) +
  geom_point(aes(color = species)) +
  scale_color_manual(values = c("darkred", "royalblue")) +
  labs(x = "Canopy Length (cm)",
       y = "Number of Green Leaves",
       color = "Species")
```
Though there is some overlap in the center area of the points, there is a clear trend of *Serenoa repens* having more green leaves than *Sabal etonia*. Length is not as clearly differentiated, but there is a slight trend of the length of *Serenoa repens* being shorter than that of *Sabal etonia*. These two variables could be helpful in a BLR.

#### Canopy Height vs Canopy Width by Species

```{r}
ggplot(data = palmetto,
       aes(x = width, y = height)) +
  geom_point(aes(color = species)) +
  scale_color_manual(values = c("darkred", "royalblue")) +
  labs(x = "Canopy Width (cm)",
       y = "Canopy Height (cm",
       color = "Species")
```
Canopy height and width are not clearly differentiated between species, so they may  not add significant value to a prediction model based on BLR.

### Binary Logistic Regression

Probability of a plant being either *Serenoa repens* or *Sabal etonia*

#### Formula and BLR Model 1
```{r}

f1 <- species ~ green_lvs + length + height

palm_blr1 <- glm(formula = f1,
                 data = palmetto,
                 family = 'binomial')
```

Results

```{r}

summary(palm_blr1)  ## Serenoa repens is reference lvl

blr1_tidy <- broom::tidy(palm_blr1)

blr1_fitted <- palm_blr1 %>% 
  broom::augment(type.predict = 'response')  # AIC = 5582
```

Plot probability

```{r}

ggplot(data = blr1_fitted,
       aes(x = green_lvs, y = .fitted)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species), se = FALSE) +
  labs(x = "Number of Green Leaves", 
       y = "Probability of outcome 'Serenoa repens'")
```

Visualize model outcomes with 'jtools::effect_plot()' 

```{r}

effect_plot(palm_blr1,
            pred = green_lvs,
            interval = TRUE,
            y.lab = "probablility of 'Serenoa repens'") # clearly delineated
                                                      # at 0 and 1

effect_plot(palm_blr1,
            pred = length,
            interval = TRUE,
            y.lab = "probablility of 'Serenoa repens'") # clearly delineated
                                                      # at 0 and 1

effect_plot(palm_blr1,
            pred = height,
            interval = TRUE,
            y.lab = "probability of 'Serenoa repens'") # not clearly                                                                  delineated at 0 and 1
```

#### Formula and BLR Model 2 

```{r}
f2 <- species ~ green_lvs + length

palm_blr2 <- glm(formula = f2,
                    data = palmetto,
                    family = 'binomial')

summary(palm_blr2)  # AIC = 5676


blr2_tidy <- broom::tidy(palm_blr2)

blr2_fitted <- palm_blr2 %>% 
  broom::augment(type.predict = 'response') 
```

```{r}

effect_plot(palm_blr2,
            pred = green_lvs,
            interval = TRUE,
            y.label = "probability of 'Serenoa repens'")

effect_plot(palm_blr2,
            pred = length,
            interval = TRUE,
            y.label = "probability of 'Serenoa repens'")
```

#### Compare with AIC and k-fold cross-validation

Corrected AIC
```{r}

aictab(list(palm_blr1, palm_blr2))

# Mod 1 AIC = 5581
# Mod 2 AIC = 5676
# delta AIC = 94 (!!!)
```

10-fold Cross-Validation

```{r}

set.seed(123)

tr_ctrl <- trainControl(method = 'repeatedcv', number = 10, repeats = 10)

## train the model
model1 <- train(f1, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model1  # 90.87% accurate

model2 <- train(f2, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model2  # 90.67% accurate
```

Cross-validation shows that both of our models are close to 91% accurate, and corrected AIC shows that our three-variable model is very significantly better than our two-variable model (delta AICc = 94). Therefore, we will utilize our three-variable model.

Use selected model

```{r}

```


### Model Evaluation

```{r}

```


### Data Citation
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
