---
title: "Task 1 - Palmetto BLR"
author: "Deanna Elliott"
date: "2/1/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
library(modelsummary)
library(kableExtra)
```

### Overview

This report explores the relationship between Florida palmetto species and their physical attributes. Binomial logistic regression models for classifying species from these attributes are tested using both corrected AIC and k-fold cross-validation. All data presented is taken from palmetto surveys at Archbold Biological Station in south-central Florida.

### Data Visualization

Read In Data

```{r}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  select(species, height:green_lvs) %>% 
  mutate(species = case_when(
    species == "1" ~ "Serenoa repens",
    species == "2" ~ "Sabal etonia")) %>% 
  mutate(species = as.factor(species)) %>% 
  drop_na()
```

```{r}
# pairs plot with GGally to see overall trends

pairplot <- palmetto %>% 
  ggpairs(aes(color = species)) # canopy length and green leaves seem most                                       differentiated by species
```

#### Green Leaves vs Canopy Length by Species

```{r}

ggplot(data = palmetto,
       aes(x = length, y = green_lvs)) +
  geom_point(aes(color = species)) +
  scale_color_manual(values = c("darkred", "royalblue")) +
  labs(x = "Canopy Length (cm)",
       y = "Number of Green Leaves",
       color = "Species")
```

Though there is some overlap in the center area of the points, there is a clear trend of *Serenoa repens* having more green leaves than *Sabal etonia*. Length is not as clearly differentiated, but there is a slight trend of the length of *Serenoa repens* being shorter than that of *Sabal etonia*. These two variables could be helpful in a BLR.

#### Canopy Height vs Canopy Width by Species

```{r}
ggplot(data = palmetto,
       aes(x = width, y = height)) +
  geom_point(aes(color = species)) +
  scale_color_manual(values = c("darkred", "royalblue")) +
  labs(x = "Canopy Width (cm)",
       y = "Canopy Height (cm)",
       color = "Species")
```

Canopy height and width are not clearly differentiated between species, so they may  not add significant value to a prediction model based on BLR.

### Binary Logistic Regression

Probability of a plant being either *Serenoa repens* or *Sabal etonia*

#### Formula and BLR Model 1
```{r}

f1 <- species ~ green_lvs + length + height

palm_blr_3var <- glm(formula = f1,
                 data = palmetto,
                 family = 'binomial')
```

Results

```{r}

summary(palm_blr_3var)  ## Serenoa repens is reference lvl

blr_3var_tidy <- broom::tidy(palm_blr_3var)

blr_3var_fitted <- palm_blr_3var %>% 
  broom::augment(type.predict = 'response')  # AIC = 5582
```

Plot probability

```{r}

ggplot(data = blr_3var_fitted,
       aes(x = green_lvs, y = .fitted)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species), se = FALSE) +
  labs(x = "Number of Green Leaves", 
       y = "Probability of outcome 'Serenoa repens'")
```

Visualize model outcomes with 'jtools::effect_plot()' 

```{r}

effect_plot(palm_blr_3var,
            pred = green_lvs,
            interval = TRUE,
            y.lab = "probablility of 'Serenoa repens'") # clearly delineated
                                                      # at 0 and 1

effect_plot(palm_blr_3var,
            pred = length,
            interval = TRUE,
            y.lab = "probablility of 'Serenoa repens'") # clearly delineated
                                                      # at 0 and 1

effect_plot(palm_blr_3var,
            pred = height,
            interval = TRUE,
            y.lab = "probability of 'Serenoa repens'") # not clearly                                                                  delineated at 0 and 1
```

#### Formula and BLR Model 2 

```{r}
f2 <- species ~ green_lvs + length

palm_blr_2var <- glm(formula = f2,
                    data = palmetto,
                    family = 'binomial')

summary(palm_blr_2var)  # AIC = 5676


blr_2var_tidy <- broom::tidy(palm_blr_2var)

blr_2var_fitted <- palm_blr_2var %>% 
  broom::augment(type.predict = 'response') 
```

Effect Plots

```{r}

effect_plot(palm_blr_2var,
            pred = green_lvs,
            interval = TRUE,
            y.label = "probability of 'Serenoa repens'")

effect_plot(palm_blr_2var,
            pred = length,
            interval = TRUE,
            y.label = "probability of 'Serenoa repens'")
```

#### Compare with AIC and k-fold cross-validation

Corrected AIC
```{r}

aictab(list(palm_blr_3var, palm_blr_2var))

# Mod 1 AIC = 5581
# Mod 2 AIC = 5676
# delta AIC = 94 (!!!)
```

10-fold Cross-Validation

```{r}

set.seed(123)

tr_ctrl <- trainControl(method = 'repeatedcv', number = 10, repeats = 10)

## train the model
mod_3var <- train(f1, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

mod_3var  # 90.87% accurate

mod_2var <- train(f2, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

mod_2var  # 90.67% accurate
```

### Model Comparison

```{r}

modelsummary(list(palm_blr_3var, palm_blr_2var), fmt = 2) %>% 
  kable_classic()
```

.

Cross-validation shows that both of our models are close to 91% accurate, and corrected AIC shows that our three-variable model is significantly better than our two-variable model (delta AICc = 94). Therefore, we will utilize our three-variable model.

### Data Citation
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
