---
title: "Task 2 - Parameter Estimation (Lizards)"
author: "Deanna Elliott"
date: "2/5/2022"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(Metrics)
library(kableExtra)
```

### Overview

This report utilizes nonlinear least squares regression to estimate parameters of a snout length to weight model for lizard populations at Jornada Basin LTER in New Mexico from 1989 to 2006. A general model was created to analyze predictions of body weight from snout-vent length of many lizard species. Then, a species-specific model was created for the same predictions on a subset of only male Western Whiptail lizards (*Cnemidophorus tigrisatus*). The two models were compared based on their RMSEs. 

### Data Wrangling

```{r}

liz <- read_csv(here("data", "lizard.csv")) %>% 
  select(spp:weight) %>% 
  drop_na()
```

### Fitted Model Plot

Create function

$$
W = a*(SVL)^b
$$

Linearize function

$$
ln(W) = b*ln(a) + b*ln(SVL)
$$

```{r}

liz_func <- function(a, SV_length, b){
  weight = a*(SV_length^b)
  return(weight)
  }
```

```{r}

# add log(length) and log(weight) to dataset

lizards <- liz %>% 
  mutate(log_length = log(SV_length),
         log_weight = log(weight))
```

```{r}

# linear regression on log columns to find parameter estimates

lin_mod <- lm(log(weight) ~ log_length, data = lizards)

a_est <- exp(lin_mod$coefficients[1])
b_est <- lin_mod$coefficients[2]

```

```{r}

liz_nls <- nls(weight ~ liz_func(a, SV_length, b),
      data = lizards,
      start = list(a = a_est, b = b_est))

summary(liz_nls) #rmse = 2.792

```

```{r}

# add predictions column to lizards dataset

liz_pred <- lizards %>% 
  mutate(predictions = predict(liz_nls, lizards, interval = "prediction"))
```

```{r}

# plot actual weights vs predicted (by sex)

ggplot(data = liz_pred) +
  geom_point(aes(x = SV_length, y = weight), color = "lightgrey") +
  geom_point(aes(x = SV_length, y = predictions, color = sex)) +
  scale_color_manual(values = c("coral", "royalblue")) +
  theme_classic() +
  labs(x = "Snout-Vent Length (mm)", y = "Body Weight (g)", color = "Sex")
```

**Figure 1. Snout-vent length vs body weight of lizards.** Snout-vent lengths for multiple species are plotted  against both actual actual observed body weights and the predicted body weights from the nonlinear least squares model.The model is color-coded by sex. The spread of actual weights is fairly close to the model at smaller snout-vent lengths, and gets wider at larger lengths.


```{r}

broom::tidy(liz_nls) %>% 
    mutate(
    p.value = scales::pvalue(p.value)) %>% 
  kable(col.names = c("Coefficient", "Estimate", "Standard Error", "Statistic", "P Value"), digits = c(0, 5, 6, 2, 3), caption = "<b>Table 1. Nonlinear least squares model estimates.</b> The estimates of our parameter coefficients, their standard, statistics, and p-value are reported for the gener llizard species model.") %>% 
  kable_classic() %>% 
  kable_styling(full_width = FALSE, position = "center")
```

### Male Western Whiptails

```{r}

mww <- lizards %>% 
  filter(spp == "CNTI",
         sex == "M") %>% 
  drop_na()

# linear regression on log columns to find parameter estimates

lin_mod_mww <- lm(log(weight) ~ log_length, data = mww)

a_est_mww <- exp(lin_mod$coefficients[1])
b_est_mww <- lin_mod$coefficients[2]
```

```{r}

mww_nls <- nls(weight ~ liz_func(a, SV_length, b),
      data = mww,
      start = list(a = a_est_mww, b = b_est_mww))

summary(mww_nls) #rmse = 3.40

mww_spp_pred <- mww %>% 
  mutate(predictions = predict(mww_nls, mww, interval = "prediction"))

mww_pred <- liz_pred %>% 
  filter(spp == "CNTI",
         sex == "M")
```

```{r}

ggplot() +
  geom_point(data = mww, aes(x = SV_length, y = weight), color = "gray") +
  geom_point(data = mww_pred, aes(x = SV_length, y = predictions, 
                                     color = "General")) +
  geom_point(data = mww_spp_pred, aes(x = SV_length, y = predictions, 
                                 color = "Species-Specific")) +
  scale_color_manual(name = "NLS Model", breaks = c("General", 
                                                    "Species-Specific"),
                                           values = c("General" = "black", 
                                                      "Species-Specific" = 
                                                        "blue")) +
  labs(x = "Snout-Vent Length (mm)",
       y = "Body Weight (g)") +
  theme_minimal()
```

**Figure 2. General vs species-specific nonlinear least squares model.** Snout-vent lengths are plotted against both actual actual observed body weights for male Western Whiptails (*C. tigrisatus*) and the predicted body weights from both the model trained on multiple lizard species and the model trained on only male Western Whiptails. 

### Takeaway

The two models are fairly similar, with the general model having an RMSE value of 2.80 and the species-specific model having an RMSE of 3.40. This tells us that the species-specific model is better at predicting the body weight of male Western Whiptail lizards from their snout-vent lengths.

### Data Citation

Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f
