---
title: "Task 1 - Palmetto BLR"
author: "Deanna Elliott"
date: "2/1/2022"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
```

### Overview

WORDS

### Data Visualization

Read In Data

```{r}
palmetto <- read_csv(here("data", "palmetto.csv"))
```

Subset? with only given variables

```{r}

```

Pairs plot with GGally

```{r}

```

Other plot

```{r}

```

Third plot?

```{r}

```


### Binary Logistic Regression

Probability of a plant being either Serenoa repens or Sabal etonia (BLR)

```{r}

```

Compare with AIC and  cross-validation

```{r}

```

Use selected model

```{r}

```


### Model Evaluation

```{r}

```


### Data Citation
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
